---
title: "Homework 1"
author: "Aysha Talakkal Cosumal - 23200212"
format: html
date: 06/17/2024
date-format: long
echo: false
---

## <span style="color: magenta;"> __INTRODUCTION__ </span>

In this document, we will explore the <span style="color: orange;"> __Orange__ </span> dataset, which is a small dataset with 35 recorded observations and 3 columns, recording the growth of orange trees.

## <span style="color: green;"> __DATASET DESCRIPTION__ </span>

```{r}
# loading the dataset
data("Orange")
```

Let us look at the first few rows of our dataset and a detailed description of the variables:

:::: {.columns}

::: {.column width="40%"}

```{r}
head(Orange,10)
```

:::

::: {.column width="60%"}

- <span style="color: skyblue;">___Tree___ </span> : ordered factor, indicates which tree measurement is being made on

- <span style="color: skyblue;">___age___ </span> : a numeric vector giving the age of the tree (days since 31st December, 1968)

- <span style="color: skyblue;">___circumference___ </span> : a numeric vector of trunk circumferences (mm)

:::


::::

Let us look at a summary of our variables, *age* and *circumference*:

```{r}
# provides summary statistics of each column except the first
summary(Orange[,-1]) 
```

\newpage

## <span style="color: green;"> __SCATTERPLOT__ </span>

Let us create a scatterplot of our data to see the relationship between our variables more clearly:

```{r}
# loading package for creating good plots
library(ggplot2)

# plot of age vs circumference, colored by Tree
ggplot(data = Orange, aes(x = age, y = circumference, color = Tree)) +
  
  # scatterplot
  geom_point() +
  
  # plot title, x- and y-axis labels
  labs(title = "Scatter Plot of Age vs Circumference for each Tree",
       x = "Age of Tree (days since 31/12/1968)",
       y = "Circumference (mm)") +
  
  # color options
  scale_color_discrete(name = "Tree") +
  theme_minimal() +
  theme(
    # Center and bolden the title
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

## <span style="color: green;"> __DESCRIPTION__ </span>

The above scatterplot provides a visual representation of the relationship between age of the tree in days (since 31st December, 1968) and its trunk circumference. This plot uses the `color` argument to differentiate between the trees, so that it is easier to infer from.

From the plot, we can clearly see that there is a positive relationship between <span style="color: teal;"> Age </span> and <span style="color: teal;"> Circumference </span>. This means that as the tree grows older, the trunk of the tree grows thicker. These findings are consistent with the known fact, that knowing the girth of a tree can tell us its age.
